\chapter{Methodology}

To gain insights into the proposed methods for researching the appliance of (ND)-Laplace for cluster algorithms we conducted experiments.
The experiment results are used to evaluate our method against other literature.
In this chapter we explain:
\begin{enumerate}

  \item Datasets
  \item Environmental setup.
  \item For each research question: Description of the different experiments.
  \item For each research question: Results.
\end{enumerate}

\section{Datasets}
For this research, we will use a synthetic dataset for all three research questions.
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
  \begin{tabular}{@{}lllll@{}}
    \toprule
    Records & Centers & Dimensions & Standard deviation & Research \\ \midrule
    200     & 4       & 2          & 0.60               & RQ 1     \\ \bottomrule
  \end{tabular}
\end{table}

Research question 3 uses a "real-world" dataset to properly assess the different dataset properties that are the subject of this research question.
\todo[inline]{Describe datasets}

\section{Environmental setup}
\todo[inline]{Describe the exact environment details}
\section{Experiment setup}
\subsection{Research question 1}
We propose several solutions for open issues based on the theoretical framework. \newline
\textbf{Choosing r: } Based, on the idea of chatzikokolakis et al. to lower the size of the radius if the place is crowded, we can do the same with clustering.
For this, we could use a metric like the standard division.
This metric increases based on clutteredness of the data, which allows us to generate a radius $r$ automatically regardless of domain.
Therefore, we depend on the configurability of epsilon entirely on privacy level $l$.
We define this metric as:
\todo[inline]{Standard deviation * 2}
\todo[inline]{Image example}

\textbf{Truncation: }
\subsection{Research question 2}
\subsection{Research question 3}
\section{Results}
\subsection{Research question 1}
\subsection{Research question 2}
\subsection{Research question 3}